---
title: "113th Senate Ideal Points"
author: "John Spaw"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using JAGS, we create an \textbf{ideal point model} for the voting preferences of the 113th US Senate given by
  $$y_{ij} \sim bernoulli(p_{ij}) \qquad probit(p_{ij}) = \alpha_j + \beta_j x_i.$$
We orient the model by assuming McConnell is to the right of Reid.
Ideal point estimates are scaled to have mean zero and unit variance. \newline
  

```{r}
library("runjags")
library("readr")
set.seed(111893)

#Data from http://tahk.us/s113.rda
#S113
#sen: info for each senator
#votes: rows = legislator, columns = bill
#pres2012: vote breakdown for 2012 presidential election
load(file = "s113.rda")
```

```{r}
#Baisc ideal point model for senator voting preference in JAGS
model_string <- "model {
    #Likelihood
    for(i in 1:n) {
      for(j in 1:J){
            y[i,j] ~ dbern(p[i,j])
            probit(p[i,j]) <- alpha[j] + beta[j]*x[i]
      }
    }
    
    #Priors
    for(i in 1:n) {
      x[i] ~ dnorm(0,1)
    }
    
    for(j in 1:J){
      alpha[j] ~ dnorm(0,0.01)
      beta[j] ~ dnorm(0,0.01)
    }
  }"

#Define controls for JAGs model fit:
monitor_string <- c("x")
jags_data <- list(y = S113$votes, 
                  n = nrow(S113$votes),
                  J = ncol(S113$votes))
```


```{r}
#Fit the basic ideal point model
#Model fit is saved after simulation is complete since fit takes a long time

#model_run_1 <- run.jags(model = model_string, monitor=monitor_string, 
#                      data = jags_data, 
#                      n.chains = 2,
#                      burnin = 4000,
#                      adapt = 1000,
#                      sample = 7500)

#saveRDS(model_run_1, "ideal_point_model_basic.rds")
```



```{r}
#Analysis of first model
#Load model from RDS file
model_1 <- readRDS("ideal_point_model_basic.rds")

#Make each iteration (rows) mean 0 var 1 (both chains)
model_1$mcmc[[1]] <- t(scale(t(model_1$mcmc[[1]])))
model_1$mcmc[[2]] <- t(scale(t(model_1$mcmc[[2]])))

#verify scaling is correct
#apply(model_1$mcmc[[1]], 1, mean)
#apply(model_1$mcmc[[2]], 1, mean)

#Compute mean for each legislator and orient McConnell to the right of Reid
ideal_points <- apply(model_1$mcmc[[1]], 2, mean)
ideal_points <- -ideal_points #put McConnell to the right of Reid

#Create dataframe and order on ideal point
ideal_point_df <- data.frame(rownames(sen), ideal_points, sen$party) 
ideal_point_df <-  ideal_point_df[order(ideal_point_df$ideal_points),] 
ideal_point_df$index <- seq(1, length(ideal_points))
names(ideal_point_df)[3] <- "party"


library(ggplot2)
ggplot(data = ideal_point_df, aes(index, ideal_points)) +
  geom_point(aes(color=party))
  #colors are strange
```

We fit a second, more complex model that accounts for a senator's party
and the democratic vote share in their state in the 2012 presidential election:


```{r}
#Add components for senator party and democratice vote share in 2012 Presidential Election
#Senator's party
sen <- S113$senators
sen[which(sen$party == "Indep"),] #Sanders and King are independent 
sen$party[c(38,96)] <- "D" #change independents to democrats ..... King and Sanders caucus with democrats
party <- ifelse(sen$party == "R", 0 , 1)

#Democratic vote share
votes <- S113$pres2012
votes$dem_vote_share <- votes$democratic/votes$total
sen <- merge(sen, votes, by = "state")
```

```{r}
#Define controls for JAGs model fit:
set.seed(111893)
library("runjags")
library("readr")

model_string_2 <- "
  model {
    #Likelihood
    for(i in 1:n) {
      for(j in 1:J){
            y[i,j] ~ dbern(p[i,j])
            probit(p[i,j]) <- alpha[j] + beta[j]*x[i]
            x[i] <- mu*party[i] + gamma*vote_share[i] + e[i]
      }
    }
  
    #Priors
    for(i in 1:n) {
      e[i] ~ dnorm(0,1)
    }
    for(j in 1:J){
      alpha[j] ~ dnorm(0,0.01)
      beta[j] ~ dnorm(0,0.01)
    }
  
    mu ~ dnorm(0,0.1)
    gamma ~ dnorm(0,0.1)
  }"


monitor_string_2 <- c("x")
jags_data_2 <- list(y = S113$votes, 
                  n = nrow(S113$votes),
                  J = ncol(S113$votes),
                  party = party,
                  vote_share = sen_new$dem_vote_share)
```


```{r}
#model_run_2 <- run.jags(model = model_string_2, monitor=monitor_string_2, 
#                      data = jags_data_2, 
#                      n.chains = 2,
#                      burnin = 4000,
#                      adapt = 1000,
#                      sample = 7500)

#Compilation error on line 7. Attempt to redefine node x[1]
#saveRDS(model_run2, "ideal_point_model_2.rds")
```














